On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-05-23 manuscript.Rmd
Minor tweaks. Version as submitted 1 March 2018.
Revision completed, as sent to coauthors on 20180523

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from regression
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to the environment
VIII. Conclusions", header=F)

@@

Stomata exert a strong control on the transpiration rate of leaves. During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
Plants face a dilemma in constructing leaves that minimize water loss while allowing the uptake of CO~2~. Since a membrane that is permeable to CO~2~ but not H~2~O has never evolved, all land plants have stomata in their leaves, which disrupt the cuticle and allow CO~2~ uptake. It is well known that stomata open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. When stomata are closed, water loss continues at a greatly diminished rate through the cuticle. After accounting for evaporative demand, this rate of water loss is expressed as the minimum conductance of a leaf. There is increasing recognition that the minimum conductance plays an important role in estimating water fluxes in plant canopies [@barnard2013], during heat waves [@kala2016], and in models of plant drought response [@blackman2016; @martinstpaul2017].

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

\
In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter. In this context, it is useful to briefly summarize the approach taken by the majority of vegetation models. Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq. 1).

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge. Our goals are to a) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq. 1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.

# II. Comparison of various definitions of minimum conductance

@@

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss cuticular conductance in Section V.

\
The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved. From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. We do not discuss the methodology of this measurement in detail, but note that care needs to be taken to yield reliable estimates (see Methods S2 for further information).

\
We quantitatively compared each of the different 'minimum' conductance (Table 1) definitions by synthesizing existing data. Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation). All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such. We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought). Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca. 150 species during the night (we selected only gas exchange based measurements from their database). Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~. However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis. Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark (further discussed in Section VI). Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks. In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime wither in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~). These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).

```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

# III. Problems with estimating minimum conductance from regression

\
The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.

\
In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated. We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied. In addition, it is difficult to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study. The correlation between slopes and intercepts is not unique to Eq. 1, but a general property of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).

```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

\
**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry). In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because there was low variation in environmental conditions (humidity, light, temperature). We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq. 1 is lower (Fig. 3c).

```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure3.png")
```

\
**Figure 3**. Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases. (a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.  Vertical lines are 95% confidence intervals. The grey line is a fitted loess smoother with 95% confidence interval. Note the wide confidence intervals, and frequent negative values. (b) Similar to (a), but using the published compilation by @miner2017. The grey line is a fitted loess smoother with 95% confidence interval. (c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e. right-hand side of the equation being fitted) is lower.

# IV. Use of minimum conductance in models

@@

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cuti~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favourable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data. When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

and using the diffusion constraint $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increases leaf transpiration, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

>>>>

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of plant desiccation
\t3. Towards a new model formulation
III. Comparison of various definitions of minimum conductance
IV. Problems with estimating Minimum conductance from regression
V. Cuticular conductance
VI. Contribution of stomata
VII. Nighttime conductance
VIII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to the environment
IX. Conclusions", header=F)

@@

During conditions where photosynthesis does not take place, such as severe drought, extreme temperature or low light, it seems advantageous for plants to close stomata and completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together constituting the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
Plants face a dilemma in constructing leaves that minimize water loss while allowing the uptake of CO~2~. Since a membrane that is permeable to CO~2~ but not H~2~O has never evolved, all land plants have stomata in their leaves, which disrupt the cuticle and allow CO~2~ uptake. It is well known that stomata open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. Stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata typically close during periods of severe water stress. When stomata are closed, water loss continues at a greatly diminished rate through the cuticle. After accounting for evaporative demand, this rate of water loss is expressed as the minimum conductance of a leaf. There is increasing recognition that the minimum conductance plays an important role in estimating water fluxes in plant canopies [@barnard2013], during heat waves [@kala2016], and in models of plant drought response [@blackman2016; @martinstpaul2017].

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest because of its simplicity is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, however proceeds much quicker (typically 0.5-2 days). This measurement also includes both surfaces of the leaf - not just the stomata-free side. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

\
In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter. Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on a different definition of the minimum conductance (g~0~): the value that should be used in models of plant water use [e.g. @barnard2013; @lombardozzi2017], and includes both g~min~ and a 'stomatal residual' due to the fact that stomata do not completely close during periods of zero photosynthesis [@dekauwe2015]. This discussion often centers around leaf conductance during the night [@lombardozzi2017], but we must also consider appropriate values for the minimum conductance during the day, for example during very high (or low) temperature, transient low light, and extreme drought. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.

\
We aim to improve the connection between the physiology of minimum conductance, empirical approaches, and model implementations by synthesizing the state of knowledge. Our goals are to a) demonstrate the need to include a non-zero minimum conductance in models of water-use efficiency and drought responses, and discuss current practices b) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.

@@

# II. Use of minimum conductance in models

@@

Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq. 1).

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero. The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but as we will show, this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is the effect of leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, and it is unclear whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018; @urban2017].

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favourable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this optimality framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data [though other mechanisms can also be employed to simulate this pattern, see @dewar2018]. When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

and using the diffusion constraint (Fick's law) $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 1 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increases leaf transpiration, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

\
**Figure 1**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Leaf transpiration (E~L~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

\
When plants are sufficiently water-stressed so that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term. A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves. Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves. @sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought. More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation. Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold. Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig. 2).

```{r fig.height=6, echo=FALSE, eval=includefigs}
plotfig("output/figure2.png")
```

\
**Figure 2**. Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants. The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues. (a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 2 or 4 mmol m^-2^ s^-1^ - all other parameters were equal. (b) Water potential in the soil and leaf as the dry-down progresses. (c) Progression of percent loss conductivity (PLC) of the xylem. Dashed line is at a PLC of 88%, indicating possible mortality.

## 3. Towards a new model formulation

$$ g_s = max \big[ max(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D) \big]$$

\
This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g. @misson2004]. Another advantage is that independent estimates of g~0~ can be used, and not via regression, which produces the undesirable correlation with estimates of g~1~. The above formulation is yet to be tested against data, but we propose this test to be done with data from drought and non-drought conditions.

# III. Comparison of various definitions of minimum conductance

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss some important aspects of cuticular conductance in Section V.

\
The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved. From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986]. Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain. In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S2). It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].

\
We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VII), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).

\
We quantitatively compared each of the different conductances by synthesizing existing data. Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation). The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.

\
All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such. Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca. 150 species during the night (we selected only gas exchange based measurements from their database). Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study. All data and code to reproduce the database and analyses are available online (see Acknowledgments).

\
The results of the quantitative comparison are presented in Fig. 3. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~. However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis. Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
